%*-------------------------------------------------------------------------*
%*                 Copyright 2013-2015 Core Physics, Inc.                  *
%*  Under terms of the contract to support CASL, there is a non-exclusive  *
%*  license for use of this work by or on behalf of the U.S. Government.   *
%*-------------------------------------------------------------------------*
%
% @version CVS $Id: verain.tex,v 1.40 2015/02/23 01:48:13 scott Exp $
%
\documentclass{report}
%
%  various packages that you may wish to activate for usage
\usepackage{graphics}
\usepackage{epsfig}
\usepackage{fancyhdr}
\usepackage{amssymb,amsmath}

\numberwithin{equation}{section}

\usepackage[pdftex,
  pdftitle={VERA Common Input},
  pdfauthor={www.casl.gov},
  pdfsubject={ },
  bookmarks, bookmarksopen, bookmarksnumbered,
  pdfstartview={FitH},
  colorlinks,linkcolor={blue},citecolor={blue},
  urlcolor={red}]{hyperref}

\DeclareGraphicsExtensions{.eps}
\DeclareGraphicsRule{.eps}{eps}{}{}

%  standard margins are 1.875in for 10pt document
%  this changes them to 1.25 inch
\addtolength{\oddsidemargin}{-0.525in}
\addtolength{\evensidemargin}{-0.525in}
\addtolength{\textwidth}{1.05in}

\setlength{\parindent}{0in}  % do not indent paragraphs, double space between
\addtolength{\parskip}{1.0\baselineskip}

%% card list - See pg. 75 in book
\newenvironment{cardlist}
 {\begin{list}{}
  {\setlength{\labelwidth}{2.0cm}
   \setlength{\leftmargin}{2.5cm}   % was 2
   \setlength{\labelsep}{0.25cm}    % was 0.5
   \setlength{\rightmargin}{2.0cm}
   \setlength{\topsep}{-0.2cm}
%  \setlength{\parsep}{0.5ex plus0.2ex minus0.5ex}
   \setlength{\itemsep}{0ex plus0.2ex} }}
{\end{list}}

%  Define page headers and footers
\pagestyle{fancy}
\newcommand{\tstamp}{\today}
\renewcommand{\chaptermark}[1]{\markboth{#1}{}}
\renewcommand{\sectionmark}[1]{\markright{#1}}
\lhead[\fancyplain{}{\thepage}]         {\fancyplain{}{\rightmark}}
\chead[\fancyplain{}{}]                 {\fancyplain{}{}}
\rhead[\fancyplain{}{\rightmark}]       {\fancyplain{VERA Common Input}{VERA Common Input}}
\lfoot[\fancyplain{}{}]                 {\fancyplain{}{}}
\cfoot[\fancyplain{\thepage}{}]         {\fancyplain{\thepage}{\thepage}}
\rfoot[\fancyplain{\tstamp} {\tstamp}]  {\fancyplain{}{}}

\begin{document}
\include{titlepage}
\tableofcontents

% \listoffigures

\chapter*{List of Acronyms}
\addcontentsline{toc}{chapter}{List of Acronyms}

\begin{description}
\item[CASL] Consortium for Advanced Simulation of Light Water Reactors
\item[BOC] Beginning of Cycle
\item[BWR] Boiling Water Reactor
\item[CFD] Computational Fluid Dynamics
\item[CILC] Crud-Induced Localized Corrosion
\item[CIPS] Crud-Induced Power Shift (also called AOA)
\item[CTF] COBRA-TF (Subchannel Code)
\item[DNB] Departure from Nucleate Boiling
\item[EFPD] Effective Full Power Days
\item[EOC] End of Cycle
\item[GWd/MT] Gigawatt-Days per Metric Ton Heavy Metal
\item[HFP] Hot Full Power
\item[HZP] Hot Zero Power
\item[LWR] Light Water Reactor
\item[MOC] Middle of Cycle
\item[MWd/MT] Megawatt-Days per Metric Ton Heavy Metal
\item[PCI] Pellet-Cladding Interaction
\item[PCM] Percent Mille (10$^{-5}$)
\item[PPM] Parts per Million (usually boron)
\item[PSU] Pennsylvania State University
\item[PWR] Pressurized Water Reactor 
\item[QA] Quality Assurance
\item[VERA] Virtual Environment for Reactor Applications

\end{description}

\vfill

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\chapter{Introduction}

\section{Introduction to CASL}
The Consortium for Advanced Simulation of Light Water Reactors (CASL) is the first
DOE Energy Innovation Hub, established in July 2010 for the purpose of providing
advanced modeling and simulation (ModSim) solutions for commercial nuclear reactors.

CASL's vision is to predict, with confidence, the performance of nuclear reactors
through comprehensive, science-based modeling and simulation technology that is
deployed and applied broadly throughout the nuclear energy industry to enhance safety,
reliability, and economics.

CASL's mission is to provide coupled, high-fidelity, usable modeling and simulation
capabilities needed to address light water reactor operational and safety performance-defining phenomena.

CASL's foundational technology products include CASL solutions and CASL ModSim Technologies.
CASL's ModSim technology, the Virtual Environment for Reactor Applications (VERA), provides
higher-fidelity results than the current industry approach by incorporating coupled physics
and science-based models, state-of-the-art numerical methods, modern computational science,
integrated uncertainty quantification (UQ) and validation against data from operating
pressurized water reactors (PWRs), single-effect experiments, and integral tests.

CASL will address, through new insights afforded by its ModSim technology, key nuclear
energy industry challenges to furthering power uprates, higher fuel burnup, and lifetime
extension while providing higher confidence in enhanced nuclear safety and this cleaner energy source.

The CASL Team is a consortium that consists of ten core partners and numerous contributing members.
The CASL organization is led by Oak Ridge National Lab, and CASL's research and development is executed
in six technical teams called Focus Areas (FA) and one integrating technical area.
This ground-breaking partnership provides unparalleled
collective institutional knowledge, nuclear science and engineering talent, computational
science leadership, and a record of LWR design and regulatory accomplishments!

More information on CASL can be found at the website \href{http://www.casl.gov}{www.casl.gov}.

%%%%%%%%%%%%
\section{VERA Core Simulator}
One component of VERA is the VERA Core Simulator (VERA-CS).  The core simulator is the specific collection
of multi-physics computer codes used to model and deplete a LWR core over multiple cycles.
Examples of the separate physics codes include cross sections, neutron transport, depletion,
thermal-hydraulics, and fuel performance.

The purpose of the core simulator is to provide data and boundary conditions to model
CASL Challenge Problems such as CIPS, CILC, DNB, and PCI analyses.

One important feature of the core simulator is that a single common input file is used to drive all
of the different physics codes\footnote{The only exception to this is for CFD codes,
which generally require a detailed CAD file to support mesh generation and perform meaningful analysis.}.
One benefit of using a single common input is that users only need to understand and be proficient with one input,
instead of having to understand multiple inputs for multiple physics codes.
Another benefit of using a single common input is that all codes work from a single
geometry description, and this reduces errors due to inconsistent geometries in different codes.

The most up-to-date version of this document resides in the VERA Git repository file
``VERAInExt/verain/docs/verain\_UM.pdf''.  
Please refer to this location for the latest version of the input manual.


%%%%%%%%%%%%
\section{Manual Organization}

This manual is organized into two parts.  

The first part, which includes Chapters~\ref{chap:user} through \ref{chap:depletion}, consists of a ``User's Manual'',
which describes how a user would set up a typical input.  This part of the manual gives
the most common input cards that a user would need and describes how to use them.
This part of the manual does not include a complete list of cards or show every available option.

The second part of the manual, Chapter~\ref{chap:cards}, is a ``Reference Manual'' and includes
a completing listing of every available input card.

The last chapter, Chapter \ref{chap:example}, gives several example input decks.  Additional input files can
also be found in the code installation directory.



%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{User Manual}
\label{chap:user}

The VERA common input is an ASCII file.
The VERA input is designed to be modular.  The input is split into separate modules
(or blocks) to describe the different geometric objects in the core and
to define specific modeling options for each of the physics codes.

Geometric objects are defined as the physical ``parts'' of the reactor core, which includes
fuel assemblies, control rod assemblies, removable burnable poison assemblies, and detectors.
By defining each geometric object as a separate block, the objects can be described
independently of each other and rely on very little global information.  The independent
descriptions make quality assurance (QA) easier and allows objects to be defined in one cycle
and be re-used in subsequent cycles without worrying about input conflicts.
Another advantage of the module approach is that it makes it easier to shuffle fuel assemblies,
and insert and withdraw ``inserts'' (such as control rods, detectors, and removable burnable
poison assemblies) into the fuel assemblies as the core configuration changes.

Additional modules/blocks are used to define modeling options and parameters for each of the physics
codes.  Separating the geometry description from the modeling options allows all of the
physics codes to share the same geometry description and also allows the same input to be
used with multiple physics codes.

The VERA input blocks are:
\begin{description}
\item[CASEID] This block contains an input title card.
\item[CORE] This block describes the core layout including core map, assembly locations,
  control rod locations, and assembly insert locations.  The CORE block contains data that does not change
  during a cycle depletion.
\item[STATE] These blocks describes reactor core operating parameters (statepoint values)
  at a particular point in time.
  Parameters include inlet temperature, pressure, power, and control rod positions.  
  STATE values can (and usually do) change at each statepoint.
\item[ASSEMBLY] These blocks contains the geometry and physical description of the nuclear fuel assemblies.
  The assembly descriptions do not include control rods, detectors, or inserts.
\item[INSERT] These blocks contain the geometry and physical description of the assembly inserts. 
   An insert is a generic term used to describe a removable burnable poison assembly or a thimble plug assembly.
\item[CONTROL] This block contains the geometry and physical description of a control rod assembly.
   A control rod assembly is similar to an assembly insert, except that it can move
   during operations.
\item[DETECTOR] This block contains the geometry and physical description of a detector string.
\item[EDITS] This block contains information on what edits the code should produce.
\item[COUPLING] This block contains parameters for coupling different physics codes together.
\end{description}

In addition to the blocks listed above, there are additional code-specific blocks that contain
options specific to each physics code.
Examples of code-specific blocks are {\bf COBRATF}, {\bf INSILICO}, and {\bf MPACT}.
Additional code-specific input blocks can be added as new physics codes are added to the core simulator.

The following sections in this chapter describe the most common concepts and features of each input block.
This section does not provide a comprehensive list of each input card or option on each card.  
Refer to Chapter~\ref{chap:cards} for a detailed listing of all input and options.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%  Section: Syntax
\section{Input Syntax}

VERA input files are text files that contain standard printable ASCII characters.
The data is organized in blocks with names and purpose as described in the
introduction. The start of a block is denoted by the block name enclosed in
square brackets, e.g. [STATE]. The file block structure is flat, so that there is
no hierarchy in the block segments. A start of a new block also implies the end of
the previous block. Currently, only the [STATE] block can have multiple instances
to describe reactor operating conditions. Other blocks are unique, so that a new
block with the same name of an existing block will overwrite the existing block data.
There is no required order of the blocks in input file, except for the [STATE] blocks,
in which each statepoint must be entered in the correct chronological order.

The blocks contain input cards that are generally organized as keyword-value pairs
or keyword-tag-value triplets where tag denotes the keyword name tag that can be
referenced in the other related commands. Keywords should not have blank spaces,
as the spaces usually imply delimiters in the card data.  A value can be a single
or list entry.
Input cards value entries can contain different data types, depending on the card format.
The data types are real numbers, integers, characters, and character strings.
String entries that include spaces should be enclosed in
single or double quote pairs.

The block, keyword and tag names are case sensitive.  Therefore, it is recommended
that users should not depend on the capitalization for differentiation between
entries in the file.

The exclamation mark, !,  is a special keyword that makes everything from it to the
end of line ignored for processing and is used for adding comments in an input file.

The keyword {\it include} can be used to insert the contents of another file into the input file.

Short commands are expected to complete within a single line. Longer commands, like input maps,
can be split across multiple lines.

An example input fragment showing blocks, comments, and cards, is shown below.
\begin{verbatim}
  ! comments start with an exclamation point

  [STATE]            ! block names are enclosed in square brackets
    power   85.0     ! cards with parameters(s)
    flow    80.0     ! cards and parameters are separated by one or more spaces

    rodbank A 228    ! cards can span more than one line
            B 228
            C 228
            D 228

  [CORE]             ! start of second block
    title  "Title must be enclosed in quotes if spaces are used"

\end{verbatim}

Several legacy Fortran codes have special characters that allow you to skip
or repeat input values.   There are no such special characters used in the VERA input.

In this manual, the convention used is that all input examples are shown in {\tt typewriter font}.
When input cards are used in the text (not in the examples), they are listed in {\it italic font}.   
All block names are listed with square brackets around them.


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%  Section: Core Description
\include{core}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%  Section: Assemblies
\include{assemblies}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%  Section: State
\include{state}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%  Chapter: Materials
\include{materials}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%  Chapter: Depletion
\include{depletion}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Input Card Descriptions} \label{chap:cards}
This chapter contains a complete listing of the available input cards.

The input for each block is given in separate subsections.

In this chapter, input cards are given in {\bf bold} text followed by the parameters on the card.
Following each input card is a description of the parameters on that card.

An example of an input card is:

{\bf card\_name} param1, param2
\begin{cardlist}
  \item[param1]  The first parameter on card (units)
  \item[param2]  The second parameter on card (units)
\end{cardlist}


\include{cards}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Examples}
\label{chap:example}

This chapter includes several input examples.  Additional examples can be found in the VERAIn Git repository.

%%%%%%%%%%%%%%%%%%
% Example 1
\section{Example 1 -- Full-Core}
The first example is a complete input for a full-core problem.
This problem is Problem 7 of the CASL Progression Benchmark Problems and
is based upon the publicly available description of the Watts Bar reactors.

More information on the CASL Progression Benchmark Problems can be found in the following CASL report:
\begin{itemize}
\item A. Godfrey, ``VERA Core Physics Benchmark Progression Problem Specifications,''
CASL Technical Report: \href{http://www.casl.gov/docs/CASL-U-2012-0131-004.pdf}{CASL-U-2012-0131-004}, August 2014.
\end{itemize}

More details on Problem 7 can be found in:
\begin{itemize}
\item ``Demonstration and Neutronics Coupled to Thermal-Hydraulics for a Full-Core Problem using VERA'',
CASL Technical Report: \href{http://www.casl.gov/docs/CASL-U-2013-0196-000.pdf}{CASL-U-2013-0196-000}, December 2013.
\end{itemize}

\include{example}

%%%%%%%%%%%%%%%%%%
% Example 2
\section{Example 2 -- Single-Assembly}
The second example is a partial input for a single-assembly with T/H feedback.
This problem is Problem 6 of the CASL Progression Benchmark Problems.  See:
\begin{itemize}
\item A. Godfrey, ``VERA Core Physics Benchmark Progression Problem Specifications,''                                      
CASL Technical Report: \href{http://www.casl.gov/docs/CASL-U-2012-0131-004.pdf}{CASL-U-2012-0131-004}, August 2014.
\end{itemize}

A single-assembly is defined by creating a core with one assembly in it, 
as described in the small-core geometry discussion in Section~\ref{sec:smallcoregeom}.

This input is also used to demonstrate the modular structure of the input.  The [ASSEMBLY],  [EDITS], [INSILICO],
[COBRATF], and [COUPLING] blocks are identical to Example Problem~1, and show how blocks
can be re-used in different input decks.  
These blocks are not included here, but can be copied directly from the first
example problem if the user wishes to run this problem.

\include{example2}

%%%%%%%%%%%%%%%%%%
% Example 3
\section{Example 3 -- 2D Lattice Geometry}
The third example is a complete input for a 2D lattice.
This problem is Problem 2A of the CASL Progression Benchmark Problems. See:
\begin{itemize}
\item A. Godfrey, ``VERA Core Physics Benchmark Progression Problem Specifications,''                                      
CASL Technical Report: \href{http://www.casl.gov/docs/CASL-U-2012-0131-004.pdf}{CASL-U-2012-0131-004}, August 2014.
\end{itemize}


A single-assembly is defined by creating a core with a one assembly in it,
as described in the small-core geometry description in Section~\ref{sec:smallcoregeom}.

The 2D lattice is defined by specifying an {\it axial} card with one level and defining 
reflective boundary conditions on the top and bottom of the core with the {\it bc\_top} and {\it bc\_bot}
input cards.

This example problem also shows how multiple assembly, insert, and control types can be
defined by using multiple {\it axial} cards in a single input block.


\include{example3}


\end{document}
